---
title: "Contrasts"
author: "Yasin Kaymaz"
date: "1/31/2019"
output:
  html_document:
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(DT)
library(tidyverse)
source("https://raw.githubusercontent.com/yasinkaymaz/cellRtools/master/main/functions.R")
source("code/functions.R")

load("~/data/all.markers.Lake2018.FC.Rdata")
load("~/data/markers.PvalbTac1_vs_GABA.Lake2018.FC.Rdata")
load("~/data/markers.In6_vs_others.Lake2018.FC.Rdata")
load("~/data/markers.In6a_vs_others.Lake2018.FC.Rdata")
load("~/data/markers.In6b_vs_others.Lake2018.FC.Rdata")
load("~/data/markers.In6a_vs_otherNeurons.Lake2018.FC.Rdata")
load("~/data/markers.In6b_vs_otherNeurons.Lake2018.FC.Rdata")

```

Here, the purpose is to identify genes that are exclusively expressed in In6a and In6b cell population in contrast against the background cells. 


    avg_logFC : log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group.
    pct.1 : The percentage of cells where the gene is detected in the first group
    pct.2 : The percentage of cells where the gene is detected in the second group
    p_val_adj : Adjusted p-value, based on bonferroni correction using all genes in the dataset.


## Differential expression of marker genes in each of the subsets:

```{r echo=FALSE}
InteractiveTable(FilterTable(allmarkers.subclass, allmarkers=T))
```

## Pvalb-Tac1 vs GABA clusters

Pvalb-Tac1 versus GABA clusters in PFC only. **Pvalb-Tac1** corresponds to In6b in Lake2018 while **GABA clusters** are In1a, In1b, In1c, In3, In4a, In4b, In6a, In7, In8.


    group1 = ("In6b")

vs

    group2 = ("In1a", "In1b", "In1c", "In3", "In4a", "In4b", "In6a", "In7", "In8")


```{r echo=FALSE}
InteractiveTable(FilterTable(markers.PvalbTac1_vs_GABA, allmarkers=F))
```

## In6 vs all others in BA10

In6 versus all remaining groups in BA10. 
                  
    group1 = ("In6a", "In6b")

vs

    group2 = ("Ast",  "End", "Mic", "Oli", "OPC", "Per","Ex1", "Ex2", "Ex3e", "Ex4", "Ex5b", "Ex6a", "Ex6b", "Ex8","In1a", "In1b", "In1c", "In3", "In4a", "In4b", "In7", "In8") 

                      
```{r echo=FALSE}
InteractiveTable(FilterTable(markers.In6_vs_others, allmarkers=F))
```

## In6a vs all others in BA10

In6a versus all remaining groups in BA10.

    group1 = ("In6a")

vs
    
    group2 = ("Ast",  "End", "Mic", "Oli", "OPC", "Per","Ex1", "Ex2", "Ex3e", "Ex4", "Ex5b", "Ex6a", "Ex6b", "Ex8","In1a", "In1b", "In1c", "In3", "In4a", "In4b", "In6b", "In7", "In8")
    

```{r echo=FALSE}
InteractiveTable(FilterTable(markers.In6a_vs_others, allmarkers=F))
```

## In6b vs all others in BA10

In6b versus all remaining groups in BA10.

    group1 = ("In6b")

vs
    
    group2 = ("Ast",  "End", "Mic", "Oli", "OPC", "Per","Ex1", "Ex2", "Ex3e", "Ex4", "Ex5b", "Ex6a", "Ex6b", "Ex8","In1a", "In1b", "In1c", "In3", "In4a", "In4b", "In6a", "In7", "In8")


```{r echo=FALSE}
InteractiveTable(FilterTable(markers.In6b_vs_others, allmarkers=F))
```


## In6a vs all other Neurons in BA10

In6a versus all remaining Neurons in BA10.

    group1 = ("In6a")

vs

    group2 = ("Ex1", "Ex2", "Ex3e", "Ex4", "Ex5b", "Ex6a", "Ex6b", "Ex8","In1a", "In1b", "In1c", "In3", "In4a", "In4b", "In6b", "In7", "In8")



```{r echo=FALSE}
InteractiveTable(FilterTable(markers.In6a_vs_otherNeurons, allmarkers=F))
```

## In6b vs all other Neurons in BA10

In6b versus all remaining Neurons in BA10.

    group1 = ("In6b")

vs

    group2 = ("Ex1", "Ex2", "Ex3e", "Ex4", "Ex5b", "Ex6a", "Ex6b", "Ex8","In1a", "In1b", "In1c", "In3", "In4a", "In4b", "In6a", "In7", "In8")


```{r echo=FALSE}
InteractiveTable(FilterTable(markers.In6b_vs_otherNeurons, allmarkers=F))
```



.

.

.

.

.

.

