---
title: "DimensionReduction_Clustering"
author: "Yasin Kaymaz"
date: "1/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(DT)
options(width = 80)
source("https://raw.githubusercontent.com/yasinkaymaz/cellRtools/master/main/functions.R")
```




```{r echo=FALSE}
load("~/data/GSE97930_FrontalCortex.seurat.Robj")
Lake2018fc <- SeuratObj
rm(SeuratObj)
```

## TSNE plots of **all Frontal Cortex** cells in the study 



These plots show the level of expression diversity among all Frontal cortex cells. Cells are colored based on cell class.

### Colored by subgroup labels

```{r echo=FALSE, fig.height=8, fig.width=10}
TSNEPlot(Lake2018fc, group.by="orig.ident",do.label = T)
```

Excitatory (Ex) and inhibitory (In) neuronal subtypes in the cortex; distinct cerebellar granule (Gran) cells and Purkinje (Purk)  neurons; and non-neuronal cells, including endothelial (End) cells,  smooth muscle cells or pericytes, astrocytes (Ast), oligodendrocytes  (Oli) and their precursor cells (OPCs), and microglia (Mic).



.

.

.

.

.

.
